---
title: "Introduction to didcf"
author: Evelina Gavrilova
date: "2025-04-02"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to didcf}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
library(didcf)
```

```{r setup}
library(didcf)
```

## Introduction
<div style="font-size: 14px;">
This vignette accompanies the package DiDCF. The algorithms comes from 
<a href="https://www.cesifo.org/en/publications/2023/working-paper/difference-difference-causal-forests-application-payroll-tax" target="_blank">*Gavrilova, Evelina, Audun Langørgen, and Floris Zoutman. "Difference-in-difference causal forests, with an application to payroll tax incidence in norway." (2025).*</a>
</div>

### How to install
Below I offer two options on how to install the package from GitHub. You can use *devtools* (which needs to be installed beforehand).

```{r}
#install.packages(devtools)
library(devtools)
install_github("evelinagz/didcf")

```

### estimate_didcf    

The function `estimate_DiDCF` takes in your data and estimates difference-in-difference causal forests. Here is how you can use the function:


```{r}
estimate_didcf(example_data$Y, 
               example_data$t_indicator[example_data$period==1], 
               model.matrix(~.,data=example_data[example_data$period==1,c("x_1","x_2")]), 
               1, 
               example_data$period, 
               example_data$unit_id) 
```
The example can be found also in the help file related to the function. The inputs to the function are the following:

1. **Y** - this is the outcome variable for all periods
2. **t_indicator** - this is a cross-sectional indicator for treatment, it should be the same in every period, here it is taken from the base period 1
3. **X** - these are the variables describing the margins of heterogeneity in the treatment effect of interest
4. **1** - this is the value of the base period in the variable denoting time
5. **period** -  variable denoting time
6. **unit_id** - variable denoting unit id


### vip

The function `vip` is a handy tool to extract a table from the variable importance feature of the estimated causal forest. The table contains variable names and their respective importance values, sorted in descending order.

```{r}
my_forest<-estimate_didcf(example_data$Y, 
               example_data$t_indicator[example_data$period==1], 
               model.matrix(~.,data=example_data[example_data$period==1,c("x_1","x_2")]), 
               1, 
               example_data$period, 
               example_data$unit_id) 
vip(my_forest[[1]])
```

### Example Data

The `example_data` dataset included in this package contains a simulated dataset tailored to the payroll example from the paper. Here is how you can load and view the dataset:

```{r}
# Load the example dataset
data(example_data, package = "didcf")

# View the dataset
head(example_data)
```
